---
title: "Homework"
output: html_document
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale(locale="English")
```

## 1

### visualize the trend of cases and deaths on the country level.
```{r}
library(dplyr)
library(lubridate)
library(tidyverse)
us <- read.csv("us.csv")
us$date <- ymd(us$date)
us %>% gather(key=kind,value="count",cases:deaths) %>%
  ggplot(aes(x=date,y=count,color=kind)) + geom_line()
us %>% gather(key=kind,value="count",cases) %>%
  ggplot(aes(x=date,y=count,color=kind)) + geom_line()
us %>% gather(key=kind,value="count",deaths) %>%
  ggplot(aes(x=date,y=count,color=kind)) + geom_line()

```

### order the states by some indicators 
### create a new indicator 'deathRate = deaths/cases'
### visualize the trend of cases and deaths on the state level.
```{r fig.height=40,fig.width=8}
states <- read.csv("us-states.csv")
df <- states %>% filter(date=="2020-04-26") %>% mutate(deathRate=round(deaths/cases,2))
# order by cases
df %>% arrange(desc(cases)) %>% head(10)

# order by deaths
df %>% arrange(desc(deaths)) %>% head(10)


# order by death rate
df %>% arrange(desc(deathRate)) %>% head(10)

states$date <- ymd(states$date)
states %>% gather(key="kind",value="count",cases:deaths) %>%
  ggplot(aes(x=date,y=count,color=kind)) + geom_line() + 
  facet_grid(scales = "free_y",rows=vars(state),cols=vars(kind))
```

### choose the top 300 counties with the most cases
### create a new indicator 'deathRate = deaths/cases'
```{r}
counties <- read.csv("us-counties.csv")
# the top 300 counties with the most cases
df <- counties %>% filter(date=="2020-04-26") %>% filter(county!="Unknown") %>%
  mutate(deathRate=round(deaths/cases,2)) %>%
  arrange(desc(cases)) %>% head(300)
df
```

### order the counties by some indicators 
```{r}
# order by cases
df %>% arrange(desc(cases))
# order by deaths
df %>% arrange(desc(deaths))
# order by the death rate
df %>% arrange(desc(deathRate))
```

### visualize the trend of cases and deaths on the county level.
```{r fig.height=80,fig.width=8}
counties$date <- ymd(counties$date)
counties %>% filter(county %in% df$county) %>% gather(key="kind",value="count",cases:deaths) %>%
  ggplot(aes(x=date,y=count,color=kind)) + geom_line() + 
  facet_grid(scales = "free_y",rows=vars(fips),cols=vars(kind))
```

## 2

###First, we have to do some scraping and cleaning.
```{r}
readData <- function(filename){
  dat <- read.csv(filename,nrows = 64)
  rows <- c("Population estimates, July 1, 2019,  (V2019)",
  "Persons under 5 years, percent", "Persons under 18 years, percent",
  "Persons 65 years and over, percent", "Female persons, percent","White alone, percent",
  "Black or African American alone, percent","Hispanic or Latino, percent",
  "High school graduate or higher, percent of persons age 25 years+, 2014-2018",
  "Bachelor's degree or higher, percent of persons age 25 years+, 2014-2018",
  "Persons  without health insurance, under age 65 years, percent",
  "Persons in poverty, percent",
  "Population per square mile, 2010")
  dat <- dat[dat$Fact %in% rows, ]
  dat <- dat[, c(1, seq(3, ncol(dat), by=2))]
  dat
}
dat1 <- readData("QuickFacts Apr-27-2020 (1).csv")
dat2 <- readData("QuickFacts Apr-27-2020 (2).csv")
dat3 <- readData("QuickFacts Apr-27-2020 (3).csv")
dat4 <- readData("QuickFacts Apr-27-2020 (4).csv")
dat5 <- readData("QuickFacts Apr-27-2020 (5).csv")
dat6 <- readData("QuickFacts Apr-27-2020 (6).csv")
dat7 <- readData("QuickFacts Apr-27-2020 (7).csv")
dat8 <- readData("QuickFacts Apr-27-2020 (8).csv")
dat9 <- readData("QuickFacts Apr-27-2020.csv")
df <- merge(dat1, dat2)
df <- merge(df, dat3)
df <- merge(df, dat4)
df <- merge(df, dat5)
df <- merge(df, dat6)
df <- merge(df, dat7)
df <- merge(df, dat8)
df <- merge(df, dat9)
dat <- df %>% gather(key=state, value=value, -Fact) %>% spread(Fact, value)
dat
```

### rename state column to merge 2 datasets 
```{r}

dat$state<-c("Alabama","Alaska","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","District of Columbia","Florida","Georgia","Hawaii","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")
dat

```

### filter the row by date=="2020-04-26"
```{r}
states <- states %>% filter(date=="2020-04-26")
states

```

### merge 2 datasets
### craete new indicators infectionRate, rate and toto death rate
```{r}
df <- merge(states, dat)
df$pop <- gsub(",", "", df$`Population estimates, July 1, 2019,  (V2019)`)
df$pop <- as.numeric(df$pop)
df$infectionRate <- df$cases/df$pop
df$deathRate <- df$deaths/df$pop
df$rate <- df$deaths/df$cases
df
```

### order the staes by some indicators
```{r}
df %>% arrange(desc(`Bachelor's degree or higher, percent of persons age 25 years+, 2014-2018`))
df %>% arrange(desc(`infectionRate`))
df %>% arrange(desc(`deathRate`))
```

### make the data in the data frame numerical fo further work
```{r}
df$under5 <- as.numeric(gsub("%","", df$`Persons under 5 years, percent`))
df$black <- as.numeric(gsub("%","", df$`Black or African American alone, percent`))
df$female <- as.numeric(gsub("%","", df$`Female persons, percent`))
df$highschool <- as.numeric(gsub("%","", df$`High school graduate or higher, percent of persons age 25 years+, 2014-2018`))
df$his <- as.numeric(gsub("%","", df$`Hispanic or Latino, percent`))
df$withoutInsurance <- as.numeric(gsub("%","", df$`Persons  without health insurance, under age 65 years, percent`))
df$over65 <- as.numeric(gsub("%","", df$`Persons 65 years and over, percent`))
df$poverty <- as.numeric(gsub("%","", df$`Persons in poverty, percent`))
df$under18 <- as.numeric(gsub("%","", df$`Persons under 18 years, percent`))
df$popPerMile <- as.numeric(gsub("%","", df$`Population per square mile, 2010`))
df$white <- as.numeric(gsub("%","", df$`White alone, percent`))
df$bachelor <- as.numeric(gsub("%","", df$`Bachelor's degree or higher, percent of persons age 25 years+, 2014-2018`))
```

### (1)
```{r}
df2 <- df %>% filter(date=="2020-04-26")
df2
```

### make plot to show the relationship between death rate and 11 other indicators
```{r}
ggplot(df2, aes(x=under5, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons under 5 years, percent", y="death Rate")

ggplot(df2, aes(x=under18, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons under 18 years, percent", y="death Rate")


ggplot(df2, aes(x=over65, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons 65 years and over, percent", y="death Rate")

ggplot(df2, aes(x=female, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Female persons, percent", y="death Rate")



ggplot(df2, aes(x=white, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="White alone, percent", y="death Rate")



ggplot(df2, aes(x=black, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Black or African American alone, percent", y="death Rate")



ggplot(df2, aes(x=his, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Hispanic or Latino, percent", y="death Rate")


ggplot(df2, aes(x=his, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Hispanic or Latino, percent", y="death Rate")


ggplot(df2, aes(x=highschool, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="High school graduate or higher, percent of persons age 25 years+, 2014-2018", y="death Rate")

ggplot(df2, aes(x=bachelor, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Bachelor's degree or higher, percent of persons age 25 years+, 2014-2018", y="death Rate")

ggplot(df2, aes(x=poverty, y=rate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons  without health insurance, under age 65 years, percent", y="death Rate")

```

### make plot to show the relationship between total death rate and 11 other indicators
```{r}
ggplot(df2, aes(x=under5, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons under 5 years, percent", y="Total death rate")

ggplot(df2, aes(x=under18, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons under 18 years, percent", y="Total death rate")


ggplot(df2, aes(x=over65, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons 65 years and over, percent", y="Total death rate")

ggplot(df2, aes(x=female, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Female persons, percent", y="Total death rate")



ggplot(df2, aes(x=white, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="White alone, percent", y="Total death rate")



ggplot(df2, aes(x=black, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Black or African American alone, percent", y="Total death rate")



ggplot(df2, aes(x=his, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Hispanic or Latino, percent", y="Total death rate")


ggplot(df2, aes(x=his, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Hispanic or Latino, percent", y="Total death rate")


ggplot(df2, aes(x=highschool, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="High school graduate or higher, percent of persons age 25 years+, 2014-2018", y="Total death rate")

ggplot(df2, aes(x=bachelor, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Bachelor's degree or higher, percent of persons age 25 years+, 2014-2018", y="Total death rate")

ggplot(df2, aes(x=poverty, y=deathRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons  without health insurance, under age 65 years, percent", y="Total death rate")
```

### make plot to show the relationship between infection rate and 11 other indicators
```{r}
ggplot(df2, aes(x=under5, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons under 5 years, percent", y="Total infection rate")

ggplot(df2, aes(x=under18, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons under 18 years, percent", y="Total infection rate")


ggplot(df2, aes(x=over65, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons 65 years and over, percent", y="Total infection rate")

ggplot(df2, aes(x=female, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Female persons, percent", y="Total infection rate")



ggplot(df2, aes(x=white, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="White alone, percent", y="Total infection rate")



ggplot(df2, aes(x=black, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Black or African American alone, percent", y="Total infection rate")



ggplot(df2, aes(x=his, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Hispanic or Latino, percent", y="Total infection rate")


ggplot(df2, aes(x=his, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Hispanic or Latino, percent", y="Total infection rate")


ggplot(df2, aes(x=highschool, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="High school graduate or higher, percent of persons age 25 years+, 2014-2018", y="Total infection rate")

ggplot(df2, aes(x=bachelor, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Bachelor's degree or higher, percent of persons age 25 years+, 2014-2018", y="Total infection rate")

ggplot(df2, aes(x=poverty, y=infectionRate)) + geom_point() + geom_smooth(se=F) +
  labs(x="Persons  without health insurance, under age 65 years, percent", y="Total infection rate")
```

